From Linear Optimization to Transmission Cost Allocation

### 2020-06-27 paper/playaround.tex
play around with tom's approach

\documentclass[11pt,twocolumn]{article}
\usepackage{graphicx}
\usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{amsmath}
\usepackage[colorlinks]{hyperref}
\usepackage[backend=biber]{biblatex}
\usepackage{eurosym}
\usepackage[dvipsnames]{xcolor} 
\usepackage{subcaption}
\usepackage{accents}
\usepackage[capitalise]{cleveref}

\addbibresource{main.bib}
\graphicspath{{figures/}}

\crefname{relation}{Rel.}{Rels.}
\creflabelformat{relation}{(#2#1#3)}
\crefname{constraint}{Constr.}{Constrs.}
\creflabelformat{constraint}{(#2#1#3)}

\setlength\parindent{8pt}

\newcommand{\ie}{\textit{i.e.} }
\newcommand{\eg}{\textit{e.g.} }

\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\newcommand{\note}[1]{\textcolor{Orange}{#1}}
\newcommand{\vpad}{\vspace{1mm}}
\newcommand{\hpad}{\hspace{15pt}}

\newcommand{\generation}[1][n]{g_{#1,s,t}}
\newcommand{\generationpotential}{\bar{g}_{n,s,t}}
\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
\newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
\newcommand{\capacityGeneration}{G_{n,s}}
\newcommand{\capacityGenerationUpper}{\bar{G}_{n,s}}
\newcommand{\capacityGenerationLower}{\ubar{G}_{n,s}}
\newcommand{\capacityFlow}{F_{\ell}}
\newcommand{\capacityFlowUpper}{\bar{F}_{\ell}}
\newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
\newcommand{\capexGeneration}{c_{n,s}}
\newcommand{\capexFlow}{c_{\ell}}
\newcommand{\opexGeneration}[1][n]{o_{#1,s}}
\newcommand{\demand}[1][n]{d_{#1,a,t}}
\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
\newcommand{\utility}{U_{n,a,t}}
\newcommand{\incidence}[1][n]{K_{#1,\ell}}
\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
\newcommand{\ptdfEqual}[1][n]{\ptdf[#1]^\circ}
\newcommand{\slackflow}{k_{\ell}}
\newcommand{\slack}[1][n]{k_{#1}}
\newcommand{\slackk}[1][n]{k^*_{#1}}
\newcommand{\Slack}{k_{m,n}}
\newcommand{\Slackk}{k^*_{m,n}}
\newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
\newcommand{\muuppergeneration}[1][n]{\bar{\mu}_{#1,s,t}}
\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
\newcommand{\flow}{f_{\ell,t}}
\newcommand{\cycle}{C_{\ell,c}}
\newcommand{\impedance}{x_\ell}
\newcommand{\cycleprice}{\lambda_{c,t}}
\newcommand{\injection}{p_{n,t}}
\newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
\newcommand{\netproduction}[1][n]{p^{+}_{#1,t}}
\newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}

\newcommand{\totalnetconsumption}{p^{-}_{t}}
\newcommand{\totalnetproduction}{p^{+}_{t}}
\newcommand{\totalselfconsumption}{p^{\circ}_{t}}

@@

\newcommand{\allocatePeer}[1][m \rightarrow n]{A_{#1,t}}
\newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
\newcommand{\allocateTransaction}[1][m \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}

@@

\newcommand{\impactcapexgeneration}{\Phi_{n,s,t}}
\newcommand{\impactcapexflow}{\Phi_{\ell,t}}

%math 
\newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
\newcommand{\pdv}[2]{\frac{\partial #1}{\partial #2}}

\title{From Linear Optimization to Transmission Cost Allocation}
\author{F. Hofmann. T. Brown}

\begin{abstract}
Maximizing the welfare of all market participants within a power system is a common approach in energy system modelling. It leads to perfectly scheduled operations of generators and  
\end{abstract}

% In the following, we show how Flow Allocation (FA) can be used for decomposing Locational Marginal Prices (LMP) in a linear optimized power system. We begin by revising the fundamentals of the Power Transfer Distribition Factors (PTDF) and the role of the slack, which plays a key role for FA. This will lead us to a generalized form of peer-to-peer allocations and transmission usage allocations based on the choice of slack. In a second step, we breakdown the full Lagrangian of a cost-optimization for a network with transmission and generation capacity expansion. By taking into account sensitivities of the flow with respect to changes in the nodal generation, we are able to derive a cost allocation which directly links to FA.

% \section{Introduction}

\subsubsection*{Nomenclature}

\begin{table}[h]
	\centering
	\begin{tabular}{ll}
        $\lmp$ & Locational Market Price at bus $n$ and  \\ & time step $t$ in \euro/MW \vpad \\
        $\demand$ & Electric demand per bus $n$, demand type $a$, \\ & time step $t$ in MW  \vpad \\
        $\generation$ & Electric generation per bus $n$, carrier $s$, \\ & time step $t$  in MW \vpad \\
        $\flow$ & Active power flow on line $\ell$, \\ & time step $t$ in MW   \vpad \\
        $\opexGeneration$ & Operational cost (OPEX) in \euro/MW \vpad \\
        $\capexGeneration$ & Capital Expenditure (CAPEX) in \euro/MW \vpad \\
        $\capexFlow$  & CAPEX per transmission line $\ell$ in \euro/MW  \vpad \\
        $\capacityGeneration$ & Generation capacity in MW \vpad \\
        $\capacityFlow$ & Transmission capacity in MW \vpad \\
        $\incidence$ & Incidence matrix \vpad 
	\end{tabular}
\end{table}

\section{Economic Context}

In long-term operation and investment planning models, the total costs $\totalcost$ of a power system include all operational expenditures (OPEX) $\totalOpexGeneration$ and capital expenditures (CAPEX) $\totalCapexGeneration$ and $\totalCapexFlow$ for production  and transmission respectively, thus
\begin{align}
  \totalcost &= \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow + \text{rest}
  \label{eq:total_cost_terms}
\end{align}
In a network design with minimal $\totalcost$, the Locational Market Price (LMP) describes the marginal price for an incremental increase of electricity demand $\demand$ of any consumer $a$ at bus $n$. It is given by the derivative of the total system cost $\totalcost$ with respect to the local demand $\demand$
\begin{align}
 \lmp = \pdv{\,\totalcost}{\,\demand}
 \label{eq:total_cost_derivative}
\end{align}
This leads to a nodal pricing where over the span of optimized timesteps $t$, the system costs are totally payed back by the consumers 
\begin{align}
 \totalcost &=  \sum_{n,a,t} \lmp \, \demand
 \label{eq:total_cost_sum}
\end{align}
% 
\begin{figure}[h]
\centering
 \includegraphics[width=.8\linewidth]{figures/price_decomposition.png}
 \caption{Schematic decomposition of the Locational Market Price $\lmp$. In power system model with optimal long-term operation and planning, the total system costs $\totalcost$ split into different cost terms, \textit{i.e} OPEX and CAPEX for production and transmission and possibly other expenditures. As $\,\totalcost$ is fully payed by the consumers, each LMP $\lmp$ can be decomposed into cost terms compensating the different expenditures at timestep $t$.}
 \label{fig:price_decomposition}
\end{figure}

% 
As a direct consequence of \cref{eq:total_cost_terms,eq:total_cost_derivative,eq:total_cost_sum}, the LMP splits into different cost terms assigned to OPEX and CAPEX for production and transmission and eventual other expenditures. We schematically show this behaviour in \cref{fig:price_decomposition}. 
Extensive investigations of the LMP as done in \cite{schweppe_spot_1988} already showed this connection however leave the question open how the costs are allocated among the components. In the following we will show how the above presented cost decomposition is extended to full peer-to-peer cost allocation including all network components.

\section{Mathematical Theory}

\subsection{Power Transfer Distribition Factors and Flow Allocation}

In linear power flow models, the Power Transfer Distribition Factors (PTDF) $\ptdf$ determine the changes in the flow on line $\ell$ for one unit (typically one MW) of net power production at bus $n$. Thus for a given production $\generation$ and demand $\demand$, they directly link to the the resulting flow on each line, 
\begin{align}
 \flow  =   \sum_n \ptdf  \left( \nodalgeneration- \nodaldemand \right)    
 \label{eq:flow_from_ptdf}
\end{align}
where $\nodalgeneration = \sum_s \generation$ and $\nodaldemand = \sum_a \demand$ combine all generators $s$ and all comsumers $a$ attached to $n$.
The PTDF have a degree of freedom: The slack $\slack$ denotes the contribution of bus $n$ to balancing out total power excess or deficit in the system. It can be dedicated to one bus, a sinlge ``slackbus``, or to several or all buses. The choice of slack modifies the PTDF according to 
\begin{align}
 \ptdf\left( \slack[m]\right)  = \ptdfEqual - \sum_m \ptdfEqual[m]  \, \slack[m]
 \label{eq:ptdf_slacked}
\end{align}
where $\ptdfEqual$ denote the PTDF with equally distributed slack.
When bus $n$ injects excess power, it has to flow to the slack; when bus $n$ extract deficit power, it has to come from the slack. Summing over all ingoing and outgoing flow changes resulting from a positive injection at $n$ yields again the slack 
\begin{align}
\sum_\ell \incidence[m] \, \ptdf =  \delta_{m,n} - \slack[m] 
\label{eq:slack}
\end{align}
Note that $\delta_{m,n}$ on the right hand side represents the positive injection at $n$.
% As shown above, the choice of slack $\slack[m]$ decides on the generators and lines to which power imbalances are distributed to. 
Established flow allocation schemes [cite] haved used this degree of freedom in order to allocate power flows and exchanges to market participants. Under the assumption that consumers account for all power flows in the grid, the slack is set to $\slack^*$ such that 
\begin{align}
 \flow  = - \sum_n \ptdf\left( \slackk[m]\right) \, \nodaldemand  
 \label{eq:flow_from_demand}
\end{align}
With such a choice the flow can be reproduced from the demand-side of the system only. Now each term in the sum on the right hand side stands for the individual contribution of consumers at node $n$ to the network flow $\flow$. In other words, each nodal demand $\nodaldemand$ induces a subflow originating from the slack $\slackk$ which all together add up to $\flow$. These subflows, in turn, can be further broken down to contributions for each bus $m$ in the slack, such that we get the subflow of individual $m \rightarrow n$ relations, that is 
\begin{align}
 \allocateTransaction = \left(  \ptdfEqual[m] - \ptdfEqual \right) \slackk[m] \, \nodaldemand
 \label{eq:allocate_transaction}
\end{align}
% 
It indicates the flow on line $\ell$ coming from generators at $m$ and supplying the demand $\nodaldemand$. When summing over all sources $m$ it yields the total subflow induced by $\nodaldemand$, the same term as in the sum on the right hand side in \cref{eq:flow_from_demand}; 
% \begin{align}
% \sum_m \allocateTransaction = -\ptdf\left(\slackk[m] \right) \,  \nodaldemand
% \end{align}
when summing over all sources and sinks, it yields again the power flow, thus
\begin{align}
\flow = \sum_{m,n} \allocateTransaction
\label{eq:transaction_sum}
\end{align}
% 
As mentioned before, the consumed power $\nodaldemand$ has to come from the slack $\slackk[m]$. As proofen in \cref{sec:proof_allocate_peer}, for each peer-to-peer relation $m \rightarrow n$, the ``traded`` power $\allocatePeer$  amounts to
\begin{align}
 \allocatePeer &= \slackk[m] \, \nodaldemand 
\label{eq:allocate_peer}
\end{align}
% 
Finally, when summing over all sinks the peer-to-peer trades yield the nodal generation (see \cref{sec:proof_sum_n_allocate_peer}) 
\begin{align}
 \sum_n \allocatePeer = \nodalgeneration[m]
 \label{eq:sum_n_allocate_peer}
\end{align}
and summing over all sources yields the nodal demand 
\begin{align}
 \sum_m \allocatePeer = \nodaldemand
 \label{eq:sum_m_allocate_peer}
\end{align}
which follows from the fact that $\sum_n \slackk = 1$.
Both allocation quantities $\allocatePeer$ and $\allocateTransaction$ can be broken down to generators $s$ or consumers $a$ by multiplying with the nodal production share $\generationshare = \generation/\sum_s \generation$ and the nodal comsumer share $\demandshare = \demand/\sum_a \demand$ respectively. \\

The solution to $\slackk$ follows from combining \cref{eq:allocate_peer,eq:sum_n_allocate_peer}, which sets it to the share of the total production $\slackk = c \, \nodalgeneration$ with $c$ being defined as $c = 1/\sum_n \nodalgeneration$. That leads to the demand $\demand$ of every single consumers $a$ being supplied by all generators $s$ in the network proportional to their gross production $\generation$. However as we discuss in \cref{sec:solution_space_of_the_slack}, the solution space can be extended to an individual slack for each node $\Slackk$.

\subsection{Network Optimisation}

% Maximise $f(x_l)$, with equality constraints $g_i(x_l)$ and inequality constraints $h_j(x_l)$
% 
% \begin{align}
%  \mathcal {L}(x_l,\lambda_i, \mu_j)=f(x_l)-\sum_i \lambda_i \, g_i(x_l) - \sum_j \mu_j \, h_j(x_l)
% \end{align}
% ...
%

We linearly cost-optimize the capacity and dispatch of a simple power system.

\begin{align}
    \underset{\generation, \capacityGeneration, \capacityFlow}{\text{min}}
    \left(\sum_{n,s} \capexGeneration \capacityGeneration + \sum_{n, s, t} \opexGeneration \generation + \sum_{\ell} \capexFlow \, \capacityFlow \right) \label{eq:minisation}
\end{align}
subject to following physical constraints. 
\\

The nodal balance constraint ensures that the amount of power that flows into a bus equals the power that flows out of a bus, thus reflects the Kirchhoff Current Law (KCL)
\begin{align}
    \sum_l \incidence \, \flow  - \nodalgeneration + \nodaldemand &= 0 \resultsin{\lmp} \Forall{n,t}
    \label[constraint]{eq:nodal_balance_lin}
\end{align}
Its shadow price mirrors the Locational Marginal Prizes (LMP) $\lmp$ per bus and time step. In a power market this is the \euro/\megawatthour-price which a consumer has to pay. Note that the flow $\flow$ in \cref{eq:nodal_balance_lin} is a passive variable only, given by \cref{eq:flow_from_ptdf}.\\

The generation $\generation$ is constraint to its nominal capacity
\begin{align}
 \generation - \generationpotential \capacityGeneration  &\le 0 \resultsin{\muuppergeneration} \Forall{n,s,t} 
 \label[constraint]{eq:upper_generation_capacity_constraint}\\ 
 - \generation &\le 0 \resultsin{\mulowergeneration} \Forall{n,s,t} 
 \label[constraint]{eq:lower_generation_capacity_constraint}
 \end{align}
where $\generationpotential \in \left[ 0,1\right]$ is the capacity factor for renewable generators. The constraints yield the KKT variables $\muuppergeneration$ and $\mulowergeneration$ which due to complementary slackness,
\begin{align}
\muuppergeneration \left( \generation - \generationpotential \, \capacityGeneration \right)  &= 0  \Forall{n,s,t} 
\label{eq:complementary_slackness_upper_generation} \\
\mulowergeneration  \, \generation &= 0 \Forall{n,s,t}
\label{eq:complementary_slackness_lower_generation} 
\end{align}
are only non-zero if the corresponding constraint is binding. \\

The transmission capacity $\capacityFlow$ limits the flow $\flow$ in both directions, such that 
\begin{align}
 \flow - \capacityFlow &\le 0 \resultsin{\muupperflow} \Forall{\ell,t} 
 \label[constraint]{eq:upper_flow_capacity_constraint} \\
 - \flow - \capacityFlow &\le 0 \resultsin{\mulowerflow} \Forall{\ell,t} 
 \label[constraint]{eq:lower_flow_capacity_constraint}
\end{align}
The yielding KKT variables $\muupperflow$ and $\mulowerflow$ are only non-zero if $\flow$ is limited by the trasmission capacity in positive or negative direction, i.e. \cref{eq:upper_flow_capacity_constraint} or \cref{eq:lower_flow_capacity_constraint} are binding. The complementary slackness 
\begin{align}
 \muupperflow \left( \flow - \capacityFlow \right)  &= 0 \Forall{\ell,t}
 \label{eq:complementary_slackness_upper_flow} \\
 \mulowerflow \left( \flow - \capacityFlow \right) &=  0 \Forall{\ell,t}
 \label{eq:complementary_slackness_lower_flow} 
\end{align}
set the respective KKT for flows staying below the thermal limit to zero. 
\\

\subsection{Allocating Nodal Payments}
\begin{align}
\notag
 & \lagrangian\left(\generation, \flow, \capacityGeneration, \capacityFlow, \boldsymbol{\lambda}, \boldsymbol{\mu} \right)   =   \\  
 \notag
 & - \sum_{n,s} \capexGeneration \capacityGeneration - \sum_{n, s, t} \opexGeneration \generation - \sum_{\ell} \capexFlow \, \capacityFlow  \\
 \notag
 &- \sum_{n,t} \lmp \left(\sum_\ell \incidence \, \flow  - \sum_s \generation + \sum_a \demand  \right)  \\ 
 \notag
 &- \sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance \, \flow  \\
 \notag
 &- \sum_{n,s,t} \muuppergeneration \left( \generation - \generationpotential \capacityGeneration \right)  + \sum_{n,s,t} \mulowergeneration \generation  \\
 &- \sum_{\ell,t} \muupperflow \left( \flow - \capacityFlow \right) + \sum_{\ell,t} \mulowerflow \left( \flow + \capacityFlow \right)     
 \label{eq:full_lagrangian}
\end{align}
% 
where $\boldsymbol{\lambda} = \left\lbrace \lmp, \cycleprice \right\rbrace $ and $\boldsymbol{\mu} = \left\lbrace \muuppergeneration, \mulowergeneration, \muupperflow, \mulowerflow \right\rbrace $ denote the set of related KKT variables. The global maximum of the Lagrangian requires stationarity with respect to all variables. The stationarity of the generation capacity variable leads to 
\begin{align}
 \pdv{\lagrangian}{\capacityGeneration}  = 0 \,\, \rightarrow \,\, 
 \capexGeneration =  \sum_t \muuppergeneration \, \generationpotential  \Forall{n,s}
 \label{eq:capexGeneration_duality}
\end{align}

the stationarity of the transmission capacity to
\begin{align}
 \pdv{\lagrangian}{\capacityFlow} = 0 \,\, \rightarrow \,\, 
 \capexFlow =  \sum_t \left( \muupperflow - \mulowerflow \right) \Forall{\ell}
 \label{eq:capexFlow_duality}
\end{align}

and the stationarity of the generation to 
\begin{align}
 \pdv{\lagrangian}{\generation} &= 0 \,\, \rightarrow \,\,  
 \opexGeneration =  \lmp - \muuppergeneration + \mulowergeneration \Forall{n,s} \label{eq:opex_duality}
\end{align}
Solving \cref{eq:opex_duality} for the $\lmp$, leads to our first representation for Locational Market Price, which we will refer to as the ``Island Solution``,
\begin{align}
\lmp  =  \opexGeneration + \muuppergeneration - \mulowergeneration \Forall{n,s,t}
\label{eq:lmp1}
\end{align}
It connects the LMP directly with the local operational price and prices for the generation capacity constraint. However, we can derive a second representation for $\lmp$. Starting from the stationarity of the flow
\begin{align}
 0 =& \pdv{\lagrangian}{\flow}  \\
 \notag
 0 =& - \sum_{m,\ell,t} \lmp[m] \, \incidence[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right)  \\
 & -\sum_{\ell,c,t} \cycleprice \, \cycle \, \impedance
\end{align}
and multipying each term with the Power Transfer Distribution Factor $\ptdf$ leaves us with

\begin{align}
  0 &= - \lmp + \sum_m \lmp[m] \, \slack[m]  - \sum_{\ell} \left( \muupperflow - \mulowerflow \right) \ptdf
  \label{eq:lmp_lmp1_reduced}
\end{align}
According to \cref{eq:slack}, the first term splits into the LMP at $n$ and the LMP weighted with the slack. The final term disappears as the $\cycle \, \impedance$ is the kernel of the PTDF $\ptdf$, so $\sum_l \cycle \, \impedance \ptdf = 0$. Solving \cref{eq:lmp_lmp1_reduced} for $\lmp$ and replacing $\lmp[m]$ of the right hand side with the expression of the Island Solution in \cref{eq:lmp1} leads to

\begin{align}
\notag
\lmp =&  \sum_m \opexGeneration[m]\, \slack[m]  + \sum_m \left( \muuppergeneration[m] - \mulowergeneration[m] \right) \slack[m] \\
& -\sum_\ell \left( \muupperflow - \mulowerflow\right) \ptdf  \Forall{n,s,t} 
\label{eq:lmp2}
\end{align}
%  
% NOTE: \cref{eq:lmp2} builds a starting point for breaking down the market value (MV) 
% of generator $(m,s)$, defined as
% \begin{align}
%  MV_s = \dfrac{\sum_t \generation \, \lmp}{\sum_t \generation}
% \end{align}
In contrast to the Island Solution in \cref{eq:lmp1}, this representation connects $\lmp$ with all other prices in the system. Moreover, it must hold for any choice of $\slack[m]$ and correspondingly decomposes the LMP to operational prices $\opexGeneration[m]$ and prices for capacity bounds for generators and transmission lines. When setting the slack to $\slackk[m] = c \, \generation[m]$, the LMP decomposes to prices of generators proportional to their production plus an extra term for the transmission usage. This directly links to the flow allocation presented above. Multiplied with the nodal demand $\nodaldemand$ the left hand side of \cref{eq:lmp2} turns into the total payment of $n$ and the right hand side into the different payment allocations, 
\begin{align}
 \lmp \, \nodaldemand = \allocateOpex + \allocateCapexGeneration + \allocateCapexFlow \Forall{n,t}
 \label{eq:lmp_allocation}
\end{align}
which we define as  
\begin{align}
 \allocateOpex &= 
 \sum_{m,s} \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
\label{eq:allocate_opexGeneration}\\
 \allocateCapexGeneration &= 
 \sum_{m,s} \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
\label{eq:allocate_capexGeneration} \\
 \allocateCapexFlow &=  
 \sum_{m,\ell} \left( \muupperflow - \mulowerflow\right) \allocateTransaction  
\label{eq:allocate_capexFlow}
\end{align}
\Cref{eq:allocate_opexGeneration} denotes the payments for generators' OPEX, \cref{eq:allocate_capexGeneration} for generators' CAPEX and \cref{eq:allocate_capexFlow} for the transmission system's CAPEX respectively. All those payment allocation build up on the phycally allocated flows presented in \cref{eq:allocate_peer,eq:allocate_transaction}. Thus with the right choice of slack the power flow allocations directly tranlate into a cost allocation.
Further, the allocated payments can be considered on a more detailed level as
\begin{align}
 \allocateOpex[n \rightarrow (m,s)] &= 
 \opexGeneration[m] \, \generationshare[m] \, \allocatePeer 
\label{eq:allocate_opexGeneration_detailed}\\
 \allocateCapexGeneration[n \rightarrow (m,s)] &= 
 \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
\label{eq:allocate_capexGeneration_detailed} \\
 \allocateCapexFlow[n \rightarrow \ell] &=  
  \left( \muupperflow - \mulowerflow\right) \sum_{m} \allocateTransaction  
\label{eq:allocate_capexFlow_detailed}
\end{align}
%

Let's have a look at the allocated OPEX in \cref{eq:allocate_opexGeneration,eq:allocate_opexGeneration_detailed} first. Consumers at bus $n$ retrieve power from different generators $(m,s)$ and accordingly compensate their operational costs. The OPEX allocation behaves like P2P tradings between producers and consumers with fixed production prices. In this way, the generator $(m,s)$ retrieves the exact amount of money from consumers that it spends on the operation. In other words, all OPEX payments to generator (m,s) sum up to the total OPEX spent at (m,s), thus 
\begin{align}
\sum_{n} \allocateOpex[n \rightarrow (m,s)] = \opexGeneration \, \generation
\label{eq:no_profit_opex}
\end{align}

>>>>

\documentclass[a4paper,10pt]{article}
\usepackage{graphicx}
\usepackage[left=1.80cm, right=1.80cm, top=2.00cm, bottom=2.00cm]{geometry}
\usepackage{amsmath}
\usepackage[colorlinks]{hyperref}
\usepackage[backend=biber,style=draft]{biblatex}
\usepackage{eurosym}
\usepackage[dvipsnames]{xcolor}
\usepackage{subcaption}
\usepackage{enumitem} % for alphabetical enumeration 
\usepackage{accents}
\usepackage[capitalise]{cleveref}

%opening
\title{}
\author{Fabian Hofmann}

\begin{abstract}

\end{abstract}

% style operators
\newcommand{\ie}{\textit{i.e.} }
\newcommand{\eg}{\textit{e.g.} }
\newcommand{\ubar}[1]{\underaccent{\bar}{#1}}
\newcommand{\note}[1]{\textcolor{Orange}{#1}}
\newcommand{\vpad}{\vspace{1mm}}
\newcommand{\hpad}{\hspace{15pt}}

% general symbols
\newcommand{\generation}[1][n]{g_{#1,s,t}}
\newcommand{\generationpotential}{\bar{g}_{n,s,t}}
\newcommand{\generationshare}[1][n]{\omega_{#1,s,t}}
\newcommand{\nodalgeneration}[1][n]{g_{#1,t}}
\newcommand{\capacityGeneration}{G_{n,s}}
\newcommand{\capacityGenerationUpper}{\bar{G}_{n,s}}
\newcommand{\capacityGenerationLower}{\ubar{G}_{n,s}}
\newcommand{\capacityFlow}{F_{\ell}}
\newcommand{\capacityFlowUpper}{\bar{F}_{\ell}}
\newcommand{\capacityFlowLower}{\ubar{F}_{\ell}}
\newcommand{\capexGeneration}{c_{n,s}}
\newcommand{\capexFlow}{c_{\ell}}
\newcommand{\opexGeneration}[1][n]{o_{#1,s}}
\newcommand{\demand}[1][n]{d_{#1,a,t}}
\newcommand{\nodaldemand}[1][n]{d_{#1,t}}
\newcommand{\demandshare}[1][n]{\omega_{#1,a,t}}
\newcommand{\utility}{U_{n,a,t}}
\newcommand{\incidence}[1][n]{K_{#1,\ell}}
\newcommand{\ptdf}[1][n]{H_{\ell,#1}}
\newcommand{\ptdfEqual}[1][n]{\ptdf[#1]^\circ}
\newcommand{\slackflow}{k_{\ell}}
\newcommand{\slack}[1][n]{k_{#1}}
\newcommand{\slackk}[1][n]{k^*_{#1}}
\newcommand{\Slack}{k_{m,n}}
\newcommand{\Slackk}{k^*_{m,n}}
\newcommand{\mulowergeneration}[1][n]{\ubar{\mu}_{#1,s,t}}
\newcommand{\muuppergeneration}[1][n]{\bar{\mu}_{#1,s,t}}
\newcommand{\mulowerflow}{\ubar{\mu}_{\ell,t}}
\newcommand{\muupperflow}{\bar{\mu}_{\ell,t}}
\newcommand{\lmp}[1][n]{\lambda_{#1,t}}
\newcommand{\flow}{f_{\ell,t}}
\newcommand{\cycle}{C_{\ell,c}}
\newcommand{\impedance}{x_\ell}
\newcommand{\cycleprice}{\lambda_{c,t}}
\newcommand{\injection}{p_{n,t}}
\newcommand{\netconsumption}[1][n]{p^{-}_{#1,t}}
\newcommand{\netproduction}[1][n]{p^{+}_{#1,t}}
\newcommand{\selfconsumption}[1][n]{p^{\circ}_{#1,t}}

% matrix notation
\newcommand{\incidenceM}{K}
\newcommand{\flowM}{f}
\newcommand{\injectionM}{p}
\newcommand{\slackM}{k}
\newcommand{\DirectedIncidence}{\mathcal{K}}
\newcommand{\InverseAPInjection}{\mathcal{J}}
\newcommand\diag[1]{\operatorname{diag}\left(#1\right)}

% totals
\newcommand{\totalnetconsumption}{p^{-}_{t}}
\newcommand{\totalnetproduction}{p^{+}_{t}}
\newcommand{\totalselfconsumption}{p^{\circ}_{t}}

@@

% allocation quantities
\newcommand{\allocatePeer}[1][(m, s) \rightarrow n]{A_{#1,t}}
\newcommand{\allocateFlow}[1][n]{F_{#1,\ell,t}}
\newcommand{\allocateTransaction}[1][m \rightarrow n]{A_{#1,\ell,t}}
\newcommand{\allocateCapexGeneration}[1][n]{\mathcal{C}^{G}_{#1,t}}
\newcommand{\allocateCapexFlow}[1][n]{\mathcal{C}^{F}_{#1,t}}
\newcommand{\allocateOpex}[1][n]{\mathcal{O}_{#1,t}}
\newcommand{\allocateEmissionCost}[1][n]{\mathcal{E}_{#1,t}}

@@

%math 
\newcommand{\resultsin}[1]{\hspace{6pt} \bot  \hspace{6pt} #1}
\newcommand{\Forall}[1]{\hspace{10pt} \forall \,\, #1 }
\newcommand{\pdv}[2]{\dfrac{\partial #1}{\partial #2}}

The total cost $\totalcost$ are composed of
\begin{align}
\totalcost &= \totalOpexGeneration + \totalCapexGeneration +  \totalCapexFlow  \\
&= \sum_{n,t} \left(  \pdv{\totalOpexGeneration}{\nodaldemand} + \pdv{\totalCapexGeneration}{\nodaldemand} + \pdv{\totalCapexFlow}{\nodaldemand} \right) \nodaldemand 
\label{eq:total_cost_terms}
\end{align}
% 
The total OPEX is given by 
\begin{align}
 \totalOpexGeneration& = \sum_{m,s,t} \opexGeneration[m]\,  \generation[m] 
\end{align}
% 
The OPEX payments assigned to node $n$ at time step $t$ is given by 
% 
\begin{align}
\allocateOpex  = \pdv{\totalOpexGeneration}{\nodaldemand} \, \nodaldemand & =  \sum_{m,s,t}\opexGeneration[m] \,  \allocatePeer
\end{align}
% 
with $\allocatePeer$ being defined as the power produced by generator $(m, s)$ and consumed at node $n$
\begin{align}
 \allocatePeer = \pdv{\generation[m]}{\nodaldemand} \, \nodaldemand
\end{align}
% 
The derivative $\partial \generation[m] / \partial \nodaldemand$ is not well-defined but implies a degree of freedom. However it is constraint to allocating exactly the amount of power which is produced by generator $(m,s)$, thus 
\begin{align}
\generation[m] = \sum_n \allocatePeer 
\label{eq:solution_space_slack}
\end{align}
% 
The direct payment of node $n$ to generator $(m,s)$ is then  
% 
\begin{align}
 \allocateOpex[n \rightarrow (m,s)] &= 
\opexGeneration[m] \,  \pdv{\generation[m]}{\nodaldemand} \, \nodaldemand
\label{eq:allocate_opexGeneration_detailed}\\
\end{align}

% \begin{align}
% \allocateCapexGeneration[n \rightarrow (m,s)] &= 
% \muuppergeneration[m] \, \generationshare[m] \, \allocatePeer
% \label{eq:allocate_capexGeneration_detailed} \\
% \allocateCapexFlow[n \rightarrow \ell] &=  
% \left( \muupperflow - \mulowerflow\right) \sum_{m} \allocateTransaction  
% \label{eq:allocate_capexFlow_detailed}
% \end{align}

The CAPEX for the transmission system is given by

\begin{align}
 \totalCapexFlow & = \sum_\ell \capexFlow \capacityFlow = \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right)  \flow 
\end{align}

The individual cost terms are 
\begin{align}
\allocateCapexFlow = \pdv{\totalCapexFlow}{\nodaldemand}\nodaldemand & = \pdv{\left( \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right)  \flow \right) }{\nodaldemand} \\
&= \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \ptdf 
\end{align}

\begin{align}
 \pdv{\muupperflow}{\nodaldemand} = 0 \;\; ?
\end{align}

\begin{align}
\pdv{\lagrangian}{\demand} & = \pdv{\lagrangian}{\muupperflow} \,  \pdv{\muupperflow} {\nodaldemand} \\
&= \left( \flow - \capacityFlow \right)  \,  \pdv{\muupperflow} {\nodaldemand}
\end{align}

\begin{align}
\notag
 \pdv{\lagrangian}{\demand}
 =&    \sum_{m, s, t} \opexGeneration[m] \Slack  \\
\notag
&+ \sum_{m,t} \lmp[m] \left(\sum_\ell \incidence[m] \, \ptdf  - \Slack + \delta_{m,n}  \right)  \\ 
\notag
&+ \sum_{n,s,t} \muuppergeneration \generationshare \Slack %  - \sum_{n,s,t} \mulowergeneration \generation  
\\
&+ \sum_{\ell,t} \left( \muupperflow - \mulowerflow \right) \ptdf      
\label{eq:full_lagrangian}
\end{align}

\subsubsection*{Production CAPEX}
\begin{align}
 \totalCapexGeneration& = \sum_{n,s} \capexGeneration \capacityGeneration\\
 &= \sum_{n,t} \muuppergeneration  \generation 
\end{align}

and taking the derivative

\begin{align}
\pdv{\totalCapexGeneration}{\nodaldemand} & = \pdv{ \left(  \sum_{m,t} \muuppergeneration[m]  \generation[m] \right) }{\nodaldemand} \\
&= \sum_{m,t} \generationshare  \muuppergeneration \, \Slack
\end{align}
